---
title: "Synchronization in MEA data. Code report."
author: "F. Menéndez Conde"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

We perform ANOVA tests, in order to determine whether the sex (male - female) and type (control - diagnosed) of the subject has an influence on the synchronization of movement between the patient and the therapist. We are interested in the response of the patient to the movement of the therapist. Hence, we only consider the frames where the therapist leads the movement and the patient follows. The negative and positive correlations are considered separately. The former corresponds to the patients movement being inhibited by the therapist's, and the latter to the opposite situation, where the patient movement mirrors the therapist.    

## The database

The database _MEA_segments.csv_ was generated previously from data recorded by the MEA software applied to 19 videos of therapy sessions. Three regions of interest (roi) were considered: head, torso and legs. The videos were split in 1 minute segments, one for each roi. In each segment the maximum and minimum spearman cross-correlations were recorded (with maximal lags of 15 seconds in each direction). Only the segments where the movement of both subjects was above the median is kept.     

Upload the data and select the relevant variables for the present analysis. 

```{r}
library(dplyr)
library(rstatix)
library(ggplot2)
library(ggpubr)
library(knitr)

MEA_segments<-read.csv("MEA_segments.csv")%>%
  select(video,zone,spearman_max,lag_spearman_max,spearman_min,lag_spearman_min,sex,type)
```

The data has a total of `r nrow(MEA_segments)` observations. 
We look at ten of them, randomly chosen. 

```{r}
n<-nrow(MEA_segments)
x<-sample(n,10)
kable(MEA_segments[x, ])
```

The variables _spearman_max_ and _spearman_min_ are the extreme values attained by the spearman cross correlations for each observation. The variables _lag_spearman_max and _lag_spearman_min_ give the time gap between the movement of each subject for which the extreme values are reached. A positive sign of the lag corresponds to the therapist movement preceding the patient.   

The data is filtered, keeping only the observations where the movement is led by the therapist. Also, lags with less to 1 second are discarded. This is done in order to avoid external noise that may give spurious simultaneous movement of the subjects. 

```{r}
x1<-which(MEA_segments$spearman_max+MEA_segments$spearman_min>0&
          MEA_segments$lag_spearman_max>1)
x2<-which(MEA_segments$spearman_max+MEA_segments$spearman_min<0&
          MEA_segments$lag_spearman_min>1)

MEA_segments<-MEA_segments[c(x1,x2), ]
```

The database is now reduced to `r nrow(MEA_segments)` observations, each corresponding to a one minute segment and roi for which the therapist leads. They are distributed into the different 19 videos as follows:

```{r}
MEA_segments <- group_by(MEA_segments,video)
segments_videos <- reframe(MEA_segments,segments=n())
kable(segments_videos)
MEA_segments<-ungroup(MEA_segments)
```

## Data preparation and visualization

Look into the values of the extremal cross-correlations for each segment, and the relation between maximal and minimal. 

```{r}
ggplot(MEA_segments,aes(spearman_max,spearman_min))+geom_point()
correlation <-cor(MEA_segments$spearman_max,MEA_segments$spearman_min)
print(summary(MEA_segments$spearman_max))
print(summary(MEA_segments$spearman_min))

n1 <- length(which(MEA_segments$spearman_max<0))
n2 <- length(which(MEA_segments$spearman_min>0))
```

The correlation between the values of the maximal and minimal correlations is `r correlation`. The maximal correlations are positive for every observation but `r n1` of them. The minimal are negative for all but `r n2` of them.  

Next, we create the _Mean_corr_by_video_ database with the mean maximal and minimal cross correlations for each video. This database will be used to perform the ANOVA. 

```{r}
MEA_segments<-group_by(MEA_segments,video,sex,type)
Mean_corr_by_video<-reframe(MEA_segments,mean_pc=mean(spearman_max),mean_nc=mean(spearman_min),sd_pc=sd(spearman_max),sd_nc=sd(spearman_min))
MEA_segments<-ungroup(MEA_segments)
kable(Mean_corr_by_video)
```


The values corresponding to the relevant variables (sex and type) are shown below:

```{r}
Mean_corr_by_video<-group_by(Mean_corr_by_video,sex,type)

table_1<-reframe(Mean_corr_by_video,mean_pc=mean(mean_pc),
mean_nc=mean(mean_nc),sd_pc=mean(sd_pc),sd_nc=mean(sd_nc))
kable(table_1)
Mean_corr_by_video<-ungroup(Mean_corr_by_video)
```

The variables _mean_pc_ and _mean_nc_ are respectively the averages of the maximal (positive) and minimal (negative) cross correlations.
Analogously for the standard deviations.

Visualization:

```{r}
ggplot(Mean_corr_by_video,aes(x=sex,y=mean_pc,color=type))+geom_boxplot()+ggtitle("Distribution of mean max cc per video")
ggplot(Mean_corr_by_video,aes(x=sex,y=mean_nc,color=type))+geom_boxplot()+ggtitle("Distribution of mean min cc per video")
```

## Analysis of variance

Test for normality and homogeneity (positive correlations).

```{r}
model  <- lm(mean_pc ~ sex*type, data = Mean_corr_by_video)
ggqqplot(residuals(model))
shapiro_test(residuals(model))
plot(model,1)
Mean_corr_by_video %>% levene_test(mean_pc ~ sex*type)
```

The normality and homogeinity assumptions are not rejected. 

Test for normality and homogeneity (negative correlations).

```{r}
model  <- lm(mean_nc ~ sex*type, data = Mean_corr_by_video)
ggqqplot(residuals(model))
shapiro_test(residuals(model))
plot(model,1)
Mean_corr_by_video %>% levene_test(mean_nc ~ sex*type)
```

The normality and homogeinity assumptions are not rejected. 

Now, we look for outliers. For positive correlations:

```{r}
temp<-Mean_corr_by_video%>%group_by(sex,type)%>%identify_outliers(mean_pc)
print(temp)
```

For negative correlations:

```{r}
temp<-Mean_corr_by_video%>%group_by(sex,type)%>%identify_outliers(mean_nc)
print(temp)
```

No outliers were found in either case. 

```{r, echo=FALSE}
Mean_corr_by_video_nc<-Mean_corr_by_video
if (length(x)>0){
  Mean_corr_by_video_nc<-Mean_corr_by_video_nc[-x,]
}
Mean_corr_by_video_pc<-Mean_corr_by_video
if (length(x)>0){
  Mean_corr_by_video_pc<-Mean_corr_by_video_pc[-x,]
}
```


Perform the two way ANOVAS for _mean_pc_ and _mean_nc_

### Negative correlations
``` {r}
anova_nc <- Mean_corr_by_video %>% anova_test(mean_nc ~ sex * type)
print(anova_nc)
```

### Positive correlations
``` {r}
anova_pc <- Mean_corr_by_video %>% anova_test(mean_pc ~ sex * type)
print(anova_pc)
```

Now, we perform one way anovas filtering for sex and looking if type influences the _mean_pc_ and _mean_nc_ outcomes.

### Positive correlations, male subjects
``` {r}
res.aov <- filter(Mean_corr_by_video_pc,sex=='Male') %>% 
  anova_test(mean_pc ~ type)
print(res.aov)
```

### Positive correlations, female subjects
``` {r}
res.aov <- filter(Mean_corr_by_video_pc,sex=='Female') %>% 
  anova_test(mean_pc ~ type)
print(res.aov)
``` 

### Negative correlations, male subjects
``` {r}
res.aov <- filter(Mean_corr_by_video_nc,sex=='Male') %>% 
  anova_test(mean_nc ~ type)
print(res.aov)
```

### Negative correlations, female subjects
``` {r}
res.aov <- filter(Mean_corr_by_video_nc,sex=='Female') %>% 
  anova_test(mean_nc ~ type)
print(res.aov)
```

## Results

The two-way ANOVA showed evidence that the combination of sex and type affects the outcome when the correlations are negative ( _p<0.001_ ). Larger absolute values correspond to the male-treatment and the female-control groups. This suggests that the movement of the therapist inhibits the movement of those groups more than the other two (male-controls and female-treatment). The one-way ANOVAS showed an statistical significant difference between the female groups ( _p=0.009_ ) but not for the males 
( _p=0.122_ ). We note that this might be explained by the fact that the reduced number of male subjects.

No significant differences were found for the positive correlations ( _p=0.07_ ). However, this is not far off, and evidence might be found with further studies.