---
title: "Statistical Analysis of MEA data."
author: "Federico Menéndez Conde"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#source("MEA_functions.R")
library(dplyr)
library(ggplot2)
library(knitr)
```

Hereby, the database _MEA_segments_  is analysed. This is the list of the 30 second periods with the most movement in the MEA videos. The criterion used is for the movement in the period to be higher than the median, and that this has to hold for both subjects (patient and therapist).

```{r, echo=TRUE}
MEA_segments <- read.csv("MEA_segments.csv")%>%
  filter(mov_patient>1,mov_therapist>1)
x<-which(MEA_segments$type=="C")
y<-which(MEA_segments$type=="T")
MEA_segments$type[x] <- "Control"
MEA_segments$type[y] <- "Diagnosed"
x<-which(MEA_segments$sex=="M")
y<-which(MEA_segments$sex=="H")
MEA_segments$sex[x] <- "F"
MEA_segments$sex[y] <- "M"
MEA_segments<-mutate(MEA_segments,delta_mov=mov_therapist-mov_patient)
```

The database has `r nrow(MEA_segments)` observations of `r ncol(MEA_segments)` variables. 

Below, we show the entries of a few observations in the database.

```{r, echo=FALSE}
kable(head(MEA_segments,10))
```



```{r, echo=FALSE}
umbral <- 0.3
MEA_segments_pos<-filter(MEA_segments,correl=="positive")%>%
  filter(abs(lag_spearman_max)>0.25)%>%
  filter(spearman_max > umbral)
MEA_segments_neg<-filter(MEA_segments,correl=="negative")%>%
  filter(abs(lag_spearman_min)>0.25)%>%
  filter(spearman_min < -umbral)
```

Split the database in two: database _MEA_segments_pos_ and _MEA_segments_neg_ according with the type of correlation found (positive or negative). Remove the entries with lags small (less than a quarter of a second). Also, keep only the entries with moderate and high cross correlations (absolute values above `r umbral`).

With such criteria, there are `r nrow(MEA_segments_pos)` segments with positive correlation and `r nrow(MEA_segments_neg)` segments with negative correlation. 

An overview of the variables of interest, classified into diagnosed subjects vs control subjects.

For all the data:

```{r, echo=TRUE}
segmented_type <- group_by(MEA_segments,type)
overview_type <- summarise(segmented_type,observed=n(),mean_max_correlation=mean(spearman_max),sd_correlation=sd(spearman_max),mean_lag=mean(lag_spearman_max),sd_lag=sd(lag_spearman_max),mean_therapist_mov=mean(mov_therapist),mean_patient_mov=mean(mov_patient))
kable(overview_type)
```

For the data with positive correlations:

```{r, echo=TRUE}
segmented_type_pos <- group_by(MEA_segments_pos,type)
overview_type_pos <- summarise(segmented_type_pos,observed=n(),mean_max_correlation=mean(spearman_max),sd_correlation=sd(spearman_max),mean_lag=mean(lag_spearman_max),sd_lag=sd(lag_spearman_max),mean_therapist_mov=mean(mov_therapist),mean_patient_mov=mean(mov_patient))
kable(overview_type_pos)
```

For the data with negative correlations:

```{r, echo=TRUE}
segmented_type_neg <- group_by(MEA_segments_neg,type)
overview_type_neg <- summarise(segmented_type_neg,observed=n(),mean_max_correlation=mean(spearman_max),sd_correlation=sd(spearman_max),mean_lag=mean(lag_spearman_max),sd_lag=sd(lag_spearman_max),mean_therapist_mov=mean(mov_therapist),mean_patient_mov=mean(mov_patient))
kable(overview_type_neg)
```

Splitting into males and females.

All the data for males:


```{r, echo=TRUE}
segmented_type_m <- filter(MEA_segments,sex=="M")%>%
        group_by(type)
overview_type_m <- summarise(segmented_type_m,observed=n(),mean_max_correlation=mean(spearman_max),sd_correlation=sd(spearman_max),mean_lag=mean(lag_spearman_max),sd_lag=sd(lag_spearman_max),mean_therapist_mov=mean(mov_therapist),mean_patient_mov=mean(mov_patient))
kable(overview_type_m)
```


Males and positive correlation

```{r, echo=FALSE}
segmented_type_m_pos <- filter(MEA_segments_pos,sex=="M")%>%
        group_by(type)
overview_type_m_pos <- summarise(segmented_type_m_pos,observed=n(),mean_max_correlation=mean(spearman_max),sd_correlation=sd(spearman_max),mean_lag=mean(lag_spearman_max),sd_lag=sd(lag_spearman_max),mean_therapist_mov=mean(mov_therapist),mean_patient_mov=mean(mov_patient))
kable(overview_type_m_pos)
```


Males and negative correlation

```{r, echo=TRUE}
segmented_type_m_neg <- filter(MEA_segments_neg,sex=="M")%>%
        group_by(type)
overview_type_m_neg <- summarise(segmented_type_m_neg,observed=n(),mean_max_correlation=mean(spearman_max),sd_correlation=sd(spearman_max),mean_lag=mean(lag_spearman_max),sd_lag=sd(lag_spearman_max),mean_therapist_mov=mean(mov_therapist),mean_patient_mov=mean(mov_patient))
kable(overview_type_m_neg)
```

All the data for females:

```{r, echo=TRUE}
segmented_type_f <- filter(MEA_segments,sex=="F")%>%
        group_by(type)
overview_type_f <- summarise(segmented_type_f,observed=n(),mean_max_correlation=mean(spearman_max),sd_correlation=sd(spearman_max),mean_lag=mean(lag_spearman_max),sd_lag=sd(lag_spearman_max),mean_therapist_mov=mean(mov_therapist),mean_patient_mov=mean(mov_patient))
kable(overview_type_f)
```

Females and positive correlation

```{r, echo=TRUE}
segmented_type_f_pos <- filter(MEA_segments_pos,sex=="F")%>%
        group_by(type)
overview_type_f_pos <- summarise(segmented_type_f_pos,observed=n(),mean_max_correlation=mean(spearman_max),sd_correlation=sd(spearman_max),mean_lag=mean(lag_spearman_max),sd_lag=sd(lag_spearman_max),mean_therapist_mov=mean(mov_therapist),mean_patient_mov=mean(mov_patient))
kable(overview_type_f_pos)
```

Females and negative correlation

```{r, echo=TRUE}
segmented_type_f_neg <- filter(MEA_segments_neg,sex=="F")%>%
        group_by(type)
overview_type_f_neg <- summarise(segmented_type_f_neg,observed=n(),mean_max_correlation=mean(spearman_max),sd_correlation=sd(spearman_max),mean_lag=mean(lag_spearman_max),sd_lag=sd(lag_spearman_max),mean_therapist_mov=mean(mov_therapist),mean_patient_mov=mean(mov_patient))
kable(overview_type_f_neg)
```

Fisher Exact Test for independence for type of subject (Control/Diagnosed) versus leading subject (Patient/Therapist). 

Males. Positive correlation. Who leads.

```{r, echo=TRUE}
MEA_segments_pos_m<-filter(MEA_segments_pos,sex=="M")
table(MEA_segments_pos_m$type,MEA_segments_pos_m$leads)
#chisq.test(MEA_segments_pos_m$type,MEA_segments_pos_m$leads,)
fisher.test(MEA_segments_pos_m$type,MEA_segments_pos_m$leads)
fisher.test(MEA_segments_pos_m$type,MEA_segments_pos_m$leads,alternative="less")
fisher.test(MEA_segments_pos_m$type,MEA_segments_pos_m$leads,alternative="greater")
```


Males. Negative correlation. Who leads.
```{r, echo=TRUE}
MEA_segments_neg_m<-filter(MEA_segments_neg,sex=="M")
table(MEA_segments_neg_m$type,MEA_segments_neg_m$leads)
#chisq.test(MEA_segments_neg_m$type,MEA_segments_neg_m$leads)
fisher.test(MEA_segments_neg_m$type,MEA_segments_neg_m$leads)
fisher.test(MEA_segments_neg_m$type,MEA_segments_neg_m$leads,alternative="less")
fisher.test(MEA_segments_neg_m$type,MEA_segments_neg_m$leads,alternative="greater")
```

Females. Positive correlations. Who leads.
```{r, echo=TRUE}
MEA_segments_pos_f<-filter(MEA_segments_pos,sex=="F")
table(MEA_segments_pos_f$type,MEA_segments_pos_f$leads)
#chisq.test(MEA_segments_pos_f$type,MEA_segments_pos_f$leads,)
fisher.test(MEA_segments_pos_f$type,MEA_segments_pos_f$leads)
```

Females. Negative correlation. Who leads.
```{r, echo=TRUE}
MEA_segments_neg_f<-filter(MEA_segments_neg,sex=="M")
table(MEA_segments_neg_f$type,MEA_segments_neg_f$leads)
#chisq.test(MEA_segments_neg_f$type,MEA_segments_neg_f$leads)
fisher.test(MEA_segments_neg_f$type,MEA_segments_neg_f$leads)
fisher.test(MEA_segments_neg_f$type,MEA_segments_neg_f$leads,alternative="less")
fisher.test(MEA_segments_neg_f$type,MEA_segments_neg_f$leads,alternative="greater")

```


The next question addressed is if the type of patient (Control/Diagnosed) has an influence on whether the correlation is positive or negative. 

Males. Is the correlation positive or negative.
```{r, echo=TRUE}
MEA_segments_m<-filter(MEA_segments,sex=="M")
table(MEA_segments_m$type,MEA_segments_m$correl)
#chisq.test(MEA_segments_m$type,MEA_segments_m$correl)
fisher.test(MEA_segments_m$type,MEA_segments_m$correl)
fisher.test(MEA_segments_m$type,MEA_segments_m$correl,alternative="less")
fisher.test(MEA_segments_m$type,MEA_segments_m$correl,alternative = "greater")
```

Males when therapist leads. Is the correlation positive or negative.
```{r, echo=TRUE}
MEA_segments_m<-filter(MEA_segments,sex=="M",leads=="therapist")
table(MEA_segments_m$type,MEA_segments_m$correl)
#chisq.test(MEA_segments_m$type,MEA_segments_m$correl)
fisher.test(MEA_segments_m$correl,MEA_segments_m$type)
fisher.test(MEA_segments_m$correl,MEA_segments_m$type,alternative="less")
fisher.test(MEA_segments_m$type,MEA_segments_m$correl,alternative = "greater")
```

Males when patient leads. Is the correlation positive or negative.
```{r, echo=TRUE}
MEA_segments_m<-filter(MEA_segments,sex=="M",leads=="patient")
table(MEA_segments_m$type,MEA_segments_m$correl)
#chisq.test(MEA_segments_m$type,MEA_segments_m$correl)
fisher.test(MEA_segments_m$type,MEA_segments_m$correl)
fisher.test(MEA_segments_m$type,MEA_segments_m$correl,alternative="less")
fisher.test(MEA_segments_m$type,MEA_segments_m$correl,alternative = "greater")
```

Females. Is the correlation positive or negative.
```{r, echo=TRUE}
MEA_segments_f<-filter(MEA_segments,sex=="F")
table(MEA_segments_f$type,MEA_segments_f$correl)
#chisq.test(MEA_segments_f$type,MEA_segments_f$correl)
fisher.test(MEA_segments_f$type,MEA_segments_f$correl)
fisher.test(MEA_segments_f$type,MEA_segments_f$correl,alternative = "less")
fisher.test(MEA_segments_f$type,MEA_segments_f$correl,alternative = "greater")
```

Females when therapist leads. Is the correlation positive or negative.
```{r, echo=TRUE}
MEA_segments_f<-filter(MEA_segments,sex=="F",leads=="therapist")
table(MEA_segments_f$type,MEA_segments_f$correl)
#chisq.test(MEA_segments_f$type,MEA_segments_f$correl)
fisher.test(MEA_segments_f$type,MEA_segments_f$correl)
fisher.test(MEA_segments_f$type,MEA_segments_f$correl,alternative = "less")
fisher.test(MEA_segments_f$type,MEA_segments_f$correl,alternative = "greater")
```


Females when patient leads. Is the correlation positive or negative.
```{r, echo=TRUE}
MEA_segments_f<-filter(MEA_segments,sex=="F",leads=="patient")
table(MEA_segments_f$type,MEA_segments_f$correl)
#chisq.test(MEA_segments_f$type,MEA_segments_f$correl)
fisher.test(MEA_segments_f$type,MEA_segments_f$correl)
fisher.test(MEA_segments_f$type,MEA_segments_f$correl,alternative = "less")
fisher.test(MEA_segments_f$type,MEA_segments_f$correl,alternative = "greater")
```

Comparison of sexes when the therapist leads. 
For controls. Is the correlation positive or negative.
```{r, echo=TRUE}
MEA_segments_b<-filter(MEA_segments,type=="Control",leads=="therapist")
table(MEA_segments_b$sex,MEA_segments_b$correl)
#chisq.test(MEA_segments_f$type,MEA_segments_f$correl)
fisher.test(MEA_segments_b$sex,MEA_segments_b$correl)
fisher.test(MEA_segments_b$sex,MEA_segments_b$correl,alternative = "less")
fisher.test(MEA_segments_b$sex,MEA_segments_b$correl,alternative = "greater")
```

Comparison of sexes when the therapist leads. 
For diagnosed. Is the correlation positive or negative.
```{r, echo=TRUE}
MEA_segments_b<-filter(MEA_segments,type=="Diagnosed",leads=="therapist")
table(MEA_segments_b$sex,MEA_segments_b$correl)
#chisq.test(MEA_segments_f$type,MEA_segments_f$correl)
fisher.test(MEA_segments_b$sex,MEA_segments_b$correl)
fisher.test(MEA_segments_b$sex,MEA_segments_b$correl,alternative = "less")
fisher.test(MEA_segments_b$sex,MEA_segments_b$correl,alternative = "greater")
```


