---
title: "Database generation from the MEA data."
author: "Federico Menéndez Conde"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE,echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)
source("MEA_functions.R")
library(dplyr)
library(ggplot2)
library(knitr)
```

```{r, include=FALSE,echo=TRUE}
zones <- c("head","torso","legs")
rois <- c(paste(zones,"patient",sep="_"),paste(zones,"therapist",sep="_"))
logbook <- read.csv("logbook_first_sessions.csv",encoding = "UTF-8")
d = 1
lmax = 450
```

## Introduction

Here, we generate the database from the MEA data that will be used for the analysis. All `r ncol(logbook)` videos considered correspond to the first session with a different patient. In each case, six regions of interest (ROI) are selected, three for each subject: head, torso and legs. 


Each video is partitioned into `r 60*d` second periods. The ones with high movement of both subjects and the same ROI are registered. Cross correlations are performed and visualized for those registered periods. 


## Set-up. 


```{r completa, include=FALSE,echo=TRUE}
completa <- list()

for (v in 1:length(logbook$file)){
    video <- paste(logbook$file[v],".txt",sep="")
    first_cut <- (logbook$sampling_rate[v])*(logbook$inicial_cut[v])
    last_cut <- (logbook$sampling_rate[v])*(logbook$final_cut[v])
    x <- arma_base(video,first_cut,last_cut)
    nombre <- substr(video,1,nchar(video)-4)
    path <- paste("bases_videos_csv/",nombre,".csv",sep = "")
    write.csv(x,path,row.names = FALSE)
    completa[[logbook$file[v]]] <- x
}

temp<-list()
for(j in names(completa)){
    print(j)
    temp<-c(temp,nrow(completa[[j]]))
}
```

Basic information of the `r dim(logbook)[1]` videos is registered in table _logbook_. Here we show this table.   

```{r logbook, echo=FALSE}
kable(logbook)
```

The list _completa_ contains the data for each of the videos listed in _logbook_. As an example, _completa[['Vid12']]_ is a data set with `r nrow(completa[['Vid12']])` observations. At a sampling rate of 30 frames per second, this corresponds to 
`r round(nrow(completa[['Vid12']])/(30*60),2)` minutes of this video. We show a small fraction (90 entries, corresponding to 3 seconds) to illustrate. 

```{r, echo=FALSE}
kable(completa[['Vid12']][115:204, ],caption='Top entries for Vid12 raw data.')
```

## Exploring short lapses with high movement.

```{r echo=TRUE}
quantile = 0.7
```

Videos are partitioned into `r 60*d` second segments. Below we show the number of segments of each of the videos.

```{r}
cuantos_frames<-numeric()
for (i in names(completa)){
    partes<-length(partition_data(completa[[i]],minutes=d))
    cuantos_frames<-c(cuantos_frames,partes)
    print(partes)
}
print(c(i,": ",sum(cuantos_frames),"frames"))
```

The mean amount of movement for each roi is calculated (as recorded by the MEA software). If the means of both the therapist and patient are in the top `r 100-100*quantile` quantile, the corresponding data for that roi and time is registered in the list _lista_mov_altos_. 


```{r echo=TRUE}
lista_mov_altos <- data.frame(matrix(ncol = 15,nrow = 0))  

for (v in names(completa)){
    lista_mov_altos <- rbind(lista_mov_altos,
          crea_lista_movs(partition_data(completa[[v]],minutes=d),
                          nombre_video = v,q=quantile,lagmax=lmax))    
}

lista_mov_altos<-filter(lista_mov_altos,!leads=="-")

ids <- data.frame(id=1:nrow(lista_mov_altos))
lista_mov_altos<-cbind(ids,lista_mov_altos)

lista_mov_altos$se_mueve_mas = "-"

for (k in 1:nrow(lista_mov_altos)){
  if (lista_mov_altos$mov_patient[k]>1.1*(lista_mov_altos$mov_therapist[k])){
      lista_mov_altos$se_mueve_mas[k]="patient"
  }    
  if (lista_mov_altos$mov_patient[k]< 0.9*(lista_mov_altos$mov_therapist[k])){
      lista_mov_altos$se_mueve_mas[k]="therapist"    
  }
}
```

Adding a column that indicates whether the cross correlations are dominated by positive or by negative correlations.

```{r echo=TRUE}
lista_mov_altos$correl<-"-"
x<-which(lista_mov_altos$spearman_max>abs(lista_mov_altos$spearman_min))
y<-which(lista_mov_altos$spearman_max<abs(lista_mov_altos$spearman_min))
lista_mov_altos$correl[x]<-"positive"
lista_mov_altos$correl[y]<-"negative"
MEA_segments<-lista_mov_altos
```

Below, we show the database _MEA_segments_ containing `r nrow(MEA_segments)` segments having high mobility for both subjects. Cross correlations were evaluated and registered for each of these segments. To determine which subject is leading, the sign of the lag with maximum correlation is considered (positive when therapist leads). 

```{r, echo=TRUE}
kable(MEA_segments)
write.csv(MEA_segments,"MEA_segments.csv",row.names = FALSE)
```

The table _MEA_segments_ is saved in the file _MEA_segments.csv_  

In the following table we show the frequency in this table for each video and roi. 

```{r, echo=TRUE}
lista_mov_altos<-group_by(MEA_segments,video,zone)
kable(summarise(lista_mov_altos,frequency=n()))
```

The totals for each roi:

```{r, echo=TRUE}
lista_mov_altos<-group_by(MEA_segments,zone)
kable(summarise(lista_mov_altos,frequency=n()))
```

And for roi and control/diagnosed groups:

```{r, echo=TRUE}
lista_mov_altos<-group_by(MEA_segments,zone,type)
kable(summarise(lista_mov_altos,frequency=n()))
```

