---
title: "Transversal Analysis of MEA data."
author: "Federico Menéndez Conde"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("MEA_functions.R")
library(dplyr)
library(ggplot2)
library(knitr)
```

```{r, include=FALSE,echo=FALSE}
zones <- c("head","torso","legs")
rois <- c(paste(zones,"patient",sep="_"),paste(zones,"therapist",sep="_"))
logbook <- read.csv("logbook_first_sessions.csv",encoding = "UTF-8")
```


The MEA data obtained from videos of therapy sessions is anallysed. All `r ncol(logbook)` videos correspond to the first session with a different patient.  In each case, six regions of interest (ROI) are considered, three for each subject (patient and therapist): head, torso and legs. 


2) In the second part, the videos are partitioned into much shorter periods (30 seconds). The ones with high movement of both subjects and the same ROI are registered. Cross correlations are performed and visualized for those registered periods. 


### First part. 


```{r completa, include=FALSE}
completa <- list()

for (v in 1:length(logbook$file)){
    video <- paste(logbook$file[v],".txt",sep="")
    first_cut <- (logbook$sampling_rate[v])*(logbook$inicial_cut[v])
    last_cut <- (logbook$sampling_rate[v])*(logbook$final_cut[v])
    x <- arma_base(video,first_cut,last_cut)
    nombre <- substr(video,1,nchar(video)-4)
    path <- paste("bases_videos_csv/",nombre,".csv",sep = "")
    write.csv(x,path,row.names = FALSE)
    completa[[logbook$file[v]]] <- x
}

temp<-list()
for(j in names(completa)){
    print(j)
    temp<-c(temp,nrow(completa[[j]]))
}
print(length(temp))
```

Basic information of the `r dim(logbook)[1]` videos is registered in table _logbook_. Here we show the top entries of this table.   

```{r logbook, echo=FALSE}
cuantos=5
kable(head(logbook,cuantos))
```

The list _completa_ contains the data for each of the videos listed in _logbook_. As an example, _completa[['Vid12']]_ is a data set with `r nrow(completa[['Vid5']])` observations. At a sampling rate of 30 frames per second, this corresponds to 
`r round(nrow(completa[['Vid5']])/(30*60),2)` minutes for the video _Vid12_. Its first entries are shown below.

```{r, echo=FALSE}
kable(head(completa[['Vid12']]),caption='Top entries for Vid12 raw data.')
```



## Part 2: exploring short lapses with high movement.

```{r echo=FALSE}
d = 0.5
quantile = 0.5
```

Videos are partitioned into `r 60*d` second segments. The mean amount of movement for each roi is calculated (as recorded by the MEA software). If the means of both the therapist and patient are in the top `r 100-100*quantile` quantile, the corresponding data for that roi and time is registered in the list _lista_mov_altos_. 


```{r echo=FALSE}
lista_mov_altos <- data.frame(matrix(ncol = 15,nrow = 0))  

for (v in names(completa)){
    lista_mov_altos <- rbind(lista_mov_altos,
          crea_lista_movs(partition_data(completa[[v]],minutes=d),
                          nombre_video = v,q=quantile,lagmax=510))    
}

lista_mov_altos<-filter(lista_mov_altos,!leads=="-")

ids <- data.frame(id=1:nrow(lista_mov_altos))
lista_mov_altos<-cbind(ids,lista_mov_altos)

lista_mov_altos$se_mueve_mas = "-"

for (k in 1:nrow(lista_mov_altos)){
  if (lista_mov_altos$mov_patient[k]>1.1*(lista_mov_altos$mov_therapist[k])){
      lista_mov_altos$se_mueve_mas[k]="patient"
  }    
  if (lista_mov_altos$mov_patient[k]< 0.9*(lista_mov_altos$mov_therapist[k])){
      lista_mov_altos$se_mueve_mas[k]="therapist"    
  }
}
```


Adding a column that indicates whether the cross correlations are dominated by positive or negative correlation.

```{r echo=FALSE}
lista_mov_altos$correl<-"-"
x<-which(lista_mov_altos$spearman_max>abs(lista_mov_altos$spearman_min))
y<-which(lista_mov_altos$spearman_max<abs(lista_mov_altos$spearman_min))
lista_mov_altos$correl[x]<-"positive"
lista_mov_altos$correl[y]<-"negative"
```


Below, we show the database _lista_mov_altos_ containing `r nrow(lista_mov_altos)` segments having high mobility for both subjects. Cross correlations were evaluated and registered for each of these segments. To determine which subject is leading, the sign og the lag with maximum correlation is considered (positive when therapist leads). 

```{r, echo=FALSE}
kable(lista_mov_altos)
write.csv(lista_mov_altos,"lista_mov_altos.csv",row.names = FALSE)
```

